### 2014-08-29 paper/paper.Rmd
updated code

```{r load_packages, echo = FALSE, message = FALSE, warning = FALSE, results='hide'}
library(dplyr)
library(stringr)
library(stringr)
library(reshape2)
library(ggplot2)
library(lubridate)
library(ggvis)
library(maps)
library(ggmap)
library(rworldmap)
library (grid)    
library (scales)
library(doBy)
```

```{r load_data, echo = FALSE, message = FALSE, warning = FALSE, results='hide'}
# This is the scaffolding to import the data
sets <- c("item", "parent", "school", "scoredItem", "student")

@@

# Now add data
#student2012$name[student2012$name=="Serbia"] <- "Republic of Serbia"
#student2012$name[student2012$name=="Korea"] <- "South Korea"
#student2012$name[student2012$name=="Chinese Taipei"] <- "Taiwan"
#student2012$name[student2012$name=="Slovak Republic"] <- "Slovakia"
#student2012$name[student2012$name=="Russian Federation"] <- "Russia"
#student2012$name[student2012$name=="Perm(Russian Federation)"] <- "Russia"
#student2012$name[student2012$name=="Hong Kong-China"] <- "Hong Kong S.A.R."
#student2012$name[student2012$name=="China-Shanghai"] <- "China"
#student2012$name[student2012$name=="China-Macau"] <- "China"
#student2012$name[student2012$name=="Connecticut (USA)"] <- "United States of America"
#student2012$name[student2012$name=="Florida (USA)"] <- "United States of America"
#student2012$name[student2012$name=="Massachusetts (USA)"] <- "United States of America"
student2012$ST06Q01 <- as.numeric(student2012$ST06Q01)
student2012$PV1MATH <- as.numeric(student2012$PV1MATH)
student2012$PV1READ <- as.numeric(student2012$PV1READ)
student2012$PV1SCIE <- as.numeric(student2012$PV1SCIE)
student2012$SENWGT_STU <- as.numeric(student2012$SENWGT_STU)

school2012 = fix_country(school2012)
student2012 = fix_country(student2012)
```

@@

```
# Out of School Study time
b1 <- summaryBy(data=student2012, pvM + pvR + pvS+ ST57Q01+ ST57Q02 +ST57Q03+ ST57Q04+ ST57Q05+ ST57Q06 ~CNT,FUN=mean, na.rm=T)
names(b1) <- c("CNT","pvM","pvR","pvS","ST57Q01", "ST57Q02","ST57Q03", "ST57Q04", "ST57Q05", "ST57Q06" )
# Out of school study time (total hours HW)
b1$CNT <- reorder(factor(b1$CNT),b1$pvS, mean)
b2 <- melt(b1[,c("CNT","pvM","pvR","pvS")],id="CNT")
b2 <- cbind(b2, b1[match(b2$CNT,b1$CNT), 5:10])
qplot(CNT, value,color=variable,size=ST57Q01, data = b2)+ coord_flip()
head(b2)
qplot(ST57Q01, value,color=variable,size=ST57Q02, data = b2)+ coord_flip()+facet_wrap(~CNT)

@@

>>>>

```{r load_packages, echo = FALSE, message = FALSE, warning = FALSE, results='hide'}
library(dplyr)
library(stringr)
library(reshape2)
library(ggplot2)
library(lubridate)
library(ggvis)
library(maps)
library(ggmap)
library(rworldmap)
library (grid)    
library (scales)
library(doBy)
```

```{r load_data, echo = FALSE, message = FALSE, warning = FALSE, results='hide', cache=TRUE}
# This is the scaffolding to import the data
sets <- c("item", "parent", "school", "scoredItem", "student")

@@

# Now add data
#student2012$name[student2012$name=="Serbia"] <- "Republic of Serbia"
#student2012$name[student2012$name=="Korea"] <- "South Korea"
#student2012$name[student2012$name=="Chinese Taipei"] <- "Taiwan"
#student2012$name[student2012$name=="Slovak Republic"] <- "Slovakia"
#student2012$name[student2012$name=="Russian Federation"] <- "Russia"
#student2012$name[student2012$name=="Perm(Russian Federation)"] <- "Russia"
#student2012$name[student2012$name=="Hong Kong-China"] <- "Hong Kong S.A.R."
#student2012$name[student2012$name=="China-Shanghai"] <- "China"
#student2012$name[student2012$name=="China-Macau"] <- "China"
#student2012$name[student2012$name=="Connecticut (USA)"] <- "United States of America"
#student2012$name[student2012$name=="Florida (USA)"] <- "United States of America"
#student2012$name[student2012$name=="Massachusetts (USA)"] <- "United States of America"
student2012$ST06Q01 <- as.numeric(student2012$ST06Q01)
student2012$PV1MATH <- as.numeric(student2012$PV1MATH)
student2012$PV1READ <- as.numeric(student2012$PV1READ)
student2012$PV1SCIE <- as.numeric(student2012$PV1SCIE)
student2012$SENWGT_STU <- as.numeric(student2012$SENWGT_STU)
school2012 = fix_country(school2012)
student2012 = fix_country(student2012)
```

@@

```{r, eval=FALSE}
# Out of School Study time
b1 <- summaryBy(data=student2012, pvM + pvR + pvS+ ST57Q01+ ST57Q02 +ST57Q03+ ST57Q04+ ST57Q05+ ST57Q06 ~CNT,FUN=mean, na.rm=T)
names(b1) <- c("CNT","pvM","pvR","pvS","ST57Q01", "ST57Q02","ST57Q03", "ST57Q04", "ST57Q05", "ST57Q06" )
# Out of school study time (total hours HW)
b1$CNT <- reorder(factor(b1$CNT),b1$pvS, mean)
b2 <- melt(b1[,c("CNT","pvM","pvR","pvS")],id="CNT")
b2 <- cbind(b2, b1[match(b2$CNT,b1$CNT), 5:10])
qplot(CNT, value,color=variable,size=ST57Q01, data = b2)+ coord_flip()
head(b2)
qplot(ST57Q01, value,color=variable,size=ST57Q02, data = b2)+ coord_flip()+facet_wrap(~CNT)

@@

```{r clean, echo=FALSE}
student2012$CNT = student2012$name
student2012$OECD <- as.character(student2012$OECD)
student2012[student2012$CNT == "Albania", ]$OECD = "Albania"
student2012$math = as.numeric(student2012[, "PV1MATH"])
student2012$read = as.numeric(student2012[, "PV1READ"])
student2012$sci = as.numeric(student2012[, "PV1SCIE"])

student2012.sub = student2012 %>% dplyr::select(CNT, SCHOOLID, math, read, sci, OUTHOURS, STIDSTD, ST04Q01, ST11Q01, ST11Q02, ST15Q01, ST19Q01, ST26Q01, ST26Q02, ST26Q03, ST26Q04, ST26Q05, ST26Q06, ST26Q07, ST26Q08, ST26Q09, ST26Q10, ST26Q11, ST26Q12, ST26Q13, ST26Q14, ST27Q01, ST27Q02, ST27Q03, ST27Q04, ST27Q05, ST28Q01, OECD)

for (i in 13:26) {
    student2012.sub[, i] <- as.character(student2012.sub[, i])
    student2012.sub[is.na(student2012.sub[, i]), i] <- ""
    student2012.sub[student2012.sub[, i] == "Yes", i] <- "1"
    student2012.sub[student2012.sub[, i] == "No", i] <- "0"
    student2012.sub[, i] <- as.numeric(student2012.sub[, i])
}
for (i in 27:31) {
    student2012.sub[, i] <- as.character(student2012.sub[, i])
    student2012.sub[is.na(student2012.sub[, i]), i] <- ""
    student2012.sub[student2012.sub[, i] == "None", i] <- "0"
    student2012.sub[student2012.sub[, i] == "One", i] <- "1"
    student2012.sub[student2012.sub[, i] == "Two", i] <- "2"
    student2012.sub[student2012.sub[, i] == "Three or more", i] <- "3"
    student2012.sub[, i] <- as.numeric(student2012.sub[, i])
}
student2012.sub[, 32] <- as.character(student2012.sub[, 32])
student2012.sub[is.na(student2012.sub[,32]), 32] <- ""
student2012.sub[student2012.sub[, 32] == "0-10 books ", 32] <- "0"
student2012.sub[student2012.sub[, 32] == "11-25 books ", 32] <- "1"
student2012.sub[student2012.sub[, 32] == "26-100 books ", 32] <- "2"
student2012.sub[student2012.sub[, 32] == "101-200 books ", 32] <- "3"
student2012.sub[student2012.sub[, 32] == "201-500 books ", 32] <- "4"
student2012.sub[student2012.sub[, 32] == "More than 500 books", 32] <- "5"
student2012.sub[, 32] <- as.numeric(student2012.sub[, 32])

#aggregate possessions
student2012.sub$numposs <- apply(student2012.sub[, 13:26], 1, sum, na.rm = T)
student2012.sub$numedposs <- apply(student2012.sub[, 27:31], 1, sum, na.rm = T)
```

```{r albania-possessions, echo=FALSE}
df = melt(student2012.sub %>% group_by(CNT, numposs, OECD) %>% summarise(mmath = mean(math), msci = mean(sci), mread = mean(read)), id=c("CNT", "numposs", "OECD"))

df$group = paste0(df$CNT, df$variable)
qplot(x = numposs, y = value, data = df, facets = ~OECD, geom = "line", color = variable, group = group) + ylab("test scores") + xlab("number of possessions") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
```

```{r truancyschool, echo=FALSE}
school2012$CNT = school2012$name
school2012.sub = school2012 %>% dplyr::select(CNT, SCHOOLID, SC22Q01)
student2012.sub = student2012 %>% dplyr::select(CNT, SCHOOLID, math, read, sci, OECD)
truancy.df = melt( student2012.sub %>% 
  group_by(CNT, SCHOOLID) %>% 
  left_join(school2012.sub, by = c("CNT", "SCHOOLID")) %>% 
  group_by(CNT, SC22Q01, OECD) %>% summarise(math = mean(math), read = mean(read), sci = mean(sci)), 
  id = c("CNT", "SC22Q01", "OECD")) %>% filter(!is.na(SC22Q01))
truancy.df$group = paste0(truancy.df$CNT, truancy.df$variable)

qplot(x = SC22Q01, y = value, color = variable, facets = ~OECD, data = truancy.df, group = group, geom = "line") + xlab("reported truancy") + ylab("test scores") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
```

```{r studytime, warning = FALSE, message=FALSE, echo=FALSE}
.width = 1

data_small <- 
  student2012 %>%
  dplyr::select(OECD, CNT, OUTHOURS) %>%
  filter(!is.na(OUTHOURS))

ggplot(mapping = aes(x = OUTHOURS)) + 
  scale_x_continuous(name = "out of school study time (hours)", limits = c(0,60)) +
  scale_y_continuous(name = "density") +
  geom_histogram(
    mapping = aes(y = ..density.., fill = OECD), 
    binwidth = .width,
    data = data_small %>% filter(OECD == "Albania")
  ) + 
  scale_fill_manual(
    values = "black",
    breaks = "Albania",
    guide = guide_legend(title = "")
  ) + 
  geom_line(
    mapping = aes(group = CNT, color = OECD), 
    stat = "density", 
    alpha = 0.3,
    data = data_small %>% filter(OECD != "Albania")
  ) +
  scale_color_discrete(
    guide = guide_legend(
      title = "Other Countries",
      override.aes = list(alpha = 1)
    )

  ) +
  scale_colour_manual(values = c("#f4a582", "#92c5de")) + guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

```{r gendergap, echo=FALSE, warning = FALSE, fig.height = 10}
scores = student2012 %>%
  dplyr::select(CNT, math, read, ST04Q01, OECD) %>% 
  group_by(OECD, CNT, ST04Q01) %>%
  summarise(math = mean(math), read = mean(read))
male = scores %>% filter(ST04Q01 == "Male")
female = scores %>% filter(ST04Q01 == "Female")
scores.df = data.frame(OECD = male$OECD, CNT = male$CNT, math_difference = male$math - female$math, read_difference = male$read - female$read)

scores.df$CNT = factor(scores.df$CNT, levels = scores.df$CNT[order(scores.df$math_difference)])
scores.df$OECD = factor(scores.df$OECD, levels = scores.df$OECD[])

qplot(CNT, math_difference, data = scores.df, color = OECD, size = OECD) + coord_flip() +
  scale_y_continuous("math score gap: male - female scores", breaks=seq(-30, 30, 5)) +
  geom_hline(yintercept=0, colour="grey80") + scale_size_manual(values = c(4, 2.5, 2.5)) + xlab("") +  
  scale_colour_manual(values = c("#ca0020", "#f4a582", "#92c5de"))

scores.df$CNT = factor(scores.df$CNT, levels = scores.df$CNT[order(scores.df$read_difference)])
qplot(CNT, read_difference, data = scores.df, color = OECD, size = OECD) + coord_flip() +
  scale_y_continuous("read score gap: male - female scores") +
  geom_hline(yintercept=0, colour="grey80") + scale_size_manual(values = c(4, 2.5, 2.5)) + xlab("") +  
  scale_colour_manual(values = c("#ca0020", "#f4a582", "#92c5de"))
```
